---
title: "Prediction for Possible Daily New Cases of Corona Infection"
auther: "R.K.Mallik PhD Condidate Civil Engineering,IOE Pulchowk Campus "
output:
  html_document: default
  pdf_document: default
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

R.K.Mallik PhD Condidate Civil Engineering,IOE Pulchowk Campus

### Import the Important Library
```{r import libraries}
library(plotly)
library(TTR)
library(tidyr)
library(dplyr)
library(ggplot2)
library(forecast)
```
### Reading Data
```{r read csv file}
coronaCases<-read.csv(file = "coronaNepal1.csv")
data1<-data.frame(coronaCases)
str(data1)
y<-data1$NewCases
length(y)
total_cases<-cumsum(y)
length(total_cases)
```
### Creating Time Series File for the Observations
```{r}
corona_ts<-ts(y,frequency=1)
total_cases_ts<-ts(total_cases,frequency=1)
corona_ts
time(corona_ts)
plot(corona_ts)
acf(corona_ts)
plot(diff(corona_ts))
acf(diff(corona_ts))
```
### Creating List of date corresponding to observations
```{r}
today <- Sys.Date()
tm <- seq(0,length(y)-1, by = +1)
x <- as.Date("2020-04-1") + tm
x

```

### Ploting observations vs date in plotly package
```{r using plotly for ploting observed data}
plot_ly(x = ~x, y = ~y,
        mode = 'line',
        xlab = 'Date',
        ylab = 'No of New Cases',
        main = 'Corona Daily Infection Observations',
        text = paste(today-x, "days from today"))

plot_ly(x = ~x, y = ~cumsum(y),
        mode = 'line',
        xlab = 'Date',
        ylab = 'Cumulative Cases',
        main = 'Corona Cumulative Cases Of Infection',
        text = paste(today-x, "days from today"))
```


### Isolating Oscillating part from the observed data
```{r}
corona_ts_noise = diff(corona_ts)
z_n<-corona_ts_noise
z_n
today <- Sys.Date()
tm <- seq(1,length(y)-1, by = +1)
tm
x <- as.Date("2020-04-1") + tm
plot_ly(x = ~x, y = z_n,
        mode = 'line',
        xlab = 'Date',
        ylab = 'No of New Cases',
        main = 'noise from the daily Observations data',
        text = paste(today-x, "days from today"))
plot_ly(x = ~x, y = cumsum(z_n),
        mode = 'line',
        xlab = 'Date',
        ylab = 'Cumulative Noise Cases',
        main = 'noise from Cumulative cases of data',
        text = paste(today-x, "days from today"))
```


### Isolating Trend part from the observation
```{r}
corona_ts_trend<-corona_ts-corona_ts_noise
z_t<-corona_ts_trend
plot_ly(x = ~x, y = z_t,
        mode = 'line',
        xlab = 'Date',
        ylab = 'No of New Cases(Trend)',
        main = 'Trend part of the Daily Corona Infections',
        text = paste(today-x, "days from today"))
plot_ly(x = ~x, y = cumsum(z_t),
        mode = 'line',
        xlab = 'Date',
        ylab = 'Cumulative No of Cases(Trend)',
        main = 'Trend part of the Cumulative Corona Infections',
        text = paste(today-x, "days from today"))
```
### DECOMPOSE TIME SERIES INTO TREND AND NOISE
```{r}
#corona_ts_decompose<-decompose(corona_ts)

```
### HOLTWINTER FORECASTING 
```{r}
plot(corona_ts,main="Daily New Cases of Corona Infection")
acf(corona_ts,main="ACF of Daily New Cases")
acf(corona_ts,type="partial",main="PACF of Daily New Cases of Infection")
corona_ts_forecasts_hw <- HoltWinters(corona_ts,gamma=FALSE)
# beta= trend, gamma=Seasionality
corona_ts_forecasts_hw
corona_ts_forecasts_hw$fitted
plot(corona_ts_forecasts_hw,main='Nepal,Daily New Cases Observation vs Fitted Model',
     xlab = 'Date Count starts from April,2,2020',
     ylab ='Daily New Cases of Corona Infection')
corona_ts_forecasts_hw$SSE

# Total Cases

corona_ts_forecasts_total_cases_hw <- HoltWinters(total_cases_ts,gamma=FALSE)
corona_ts_forecasts_total_cases_hw
corona_ts_forecasts_total_cases_hw$fitted
plot(corona_ts_forecasts_total_cases_hw,main='Nepal,Cumulative Cases Observation vs Fitted Model',
     xlab = 'Date Count starts from April,2,2020',
     ylab ='Cumulative Cases of Corona Infection')
corona_ts_forecasts_total_cases_hw$SSE

```

### ONE WEEAK FORCASTED DATA
```{r}
library("forecast")
corona_ts_forecasts_hw2 <- forecast(corona_ts_forecasts_hw,h=8)
corona_ts_forecasts_hw2

plot(corona_ts_forecasts_hw2,main='Arima Model Forecast for one Week(6 June to 13 June)',
                xlab = 'Date Count starts from April,2,2020',
                ylab ='Daily New Cases of Corona Infection')
acf(corona_ts_forecasts_hw2)
# Total Cases
corona_ts_forecasts_total_cases_hw2 <- forecast(corona_ts_forecasts_total_cases_hw,h=30)
corona_ts_forecasts_total_cases_hw2
plot(corona_ts_forecasts_total_cases_hw2,main='HotWinter Model Forecast for one Week(6 June to 13 June)',
                xlab = 'Date Count starts from April,2,2020',
                ylab ='Cumulative Cases of Corona Infection')
```

### Plotting Residulas and lag
```{r}
acf(corona_ts_forecasts_hw2,lag.max=20)
plot.ts(corona_ts_forecasts_hw2$residuals)
```
### HOLT'S EXPONENTIAL SMOOTHING
```{r}
corona_ts_forecasts_exp <- HoltWinters(corona_ts,gamma=FALSE)
corona_ts_forecasts_exp
corona_ts_forecasts_exp$fitted
plot(corona_ts_forecasts_exp,main='Holts Exponential Model',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Daily New Cases of Corona Infection')

corona_ts_forecasts_exp$SSE

# Total Cases..
corona_ts_forecasts_total_cases_exp <- HoltWinters(total_cases_ts,gamma=FALSE)
corona_ts_forecasts_total_cases_exp
corona_ts_forecasts_total_cases_exp$fitted
plot(corona_ts_forecasts_total_cases_exp,main='Holts Exponential Model',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')

```
### FORCASTED DATA CORRESPONDING TO EXPONENTIAL MODEL
```{r}
corona_ts_forecasts_exp2 <- forecast(corona_ts_forecasts_exp,h=8)
corona_ts_forecasts_exp2
plot(corona_ts_forecasts_exp2,main='Exponential Model Forecast for one Week(9June to 16 June)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Daily New Cases of Corona Infection')

acf(corona_ts_forecasts_exp2,lag.max=20)
Box.test(corona_ts_forecasts_exp2$residuals,lag=20,type='Ljung-Box')
plot.ts(corona_ts_forecasts_exp2$residuals)

# Total Cases Fore cast for one week
corona_ts_forecasts_total_cases_exp2 <- forecast(corona_ts_forecasts_total_cases_exp,h=8)
corona_ts_forecasts_total_cases_exp2
plot(corona_ts_forecasts_total_cases_exp2,main='Exponential Model Forecast for one Week(9June to 16 June)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')

# Total Cases Fore cast for one month
corona_ts_forecasts_total_cases_exp2_month <- forecast(corona_ts_forecasts_total_cases_exp,h=30)
corona_ts_forecasts_total_cases_exp2_month
plot(corona_ts_forecasts_total_cases_exp2_month,main='Exponential Model Forecast for one Month(6June to 6 July)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')
# Total Cases Fore cast for two month
corona_ts_forecasts_total_cases_exp2_month <- forecast(corona_ts_forecasts_total_cases_exp,h=90)
corona_ts_forecasts_total_cases_exp2_month
plot(corona_ts_forecasts_total_cases_exp2_month,main='Exponential Model Forecast for one Week(6June to 6 September)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')

```

### EXTENDING EXPONENTIAL MODEL TO PREDICT SATURATION LEVEL
```{r}
today <- Sys.Date()
tm <- seq(0,length(y)-1, by = +1)
x <- as.Date("2020-04-1") + tm
x
length(total_cases)
data2<-data.frame(x,total_cases)
ggplot(data2,aes(x=x,y=total_cases/3000))+geom_point()+
  stat_smooth(method="glm",method.args=list(family="binomial"),se=FALSE)
#par(mar=c(4,4,1,1))
#plot(data2$x,data2$total_cases)


```



#```{r}
# Assume effective time between generations
#t = 3
# Cumulative Data Ponts after gap of t days
#total_cases
#for (i in 1:length(total_cases)){
#  total_slope[i] <- total_cases[i+1]-total_cases[i]
#}
#total_slope
#for (i in seq(length(total_cases),2,by=-t)){
#  theta[i]<-(1/t)*log(total_cases[i]/total_cases[i-t])
#  
#}
#theta
#theta1<-na.omit(theta)
#theta1
#theta2<-mean(theta1)
#theta2
# Saturation Value
#t = 1:length(total_cases)
#M<-total_cases/(1-total_cases*exp(theta1*t))
#M = 10000
#no = 1
#Nt<-M/(1+((M-no)/no)*exp(-theta2*t))
#Nt
#plot(t,Nt)
#```

## ARIMA MODEL 
```{r}
diff_corona_ts<-diff(corona_ts)
plot(diff_corona_ts)
acf(diff_corona_ts)
pacf(diff_corona_ts)
# automatic selection of p,d,andq
auto.arima(diff_corona_ts)
corona_ts_arima<-arima(diff_corona_ts,order=c(2,0,2))


library(forecast)
# Weekly Forecast
corona_ts_arima<-arima(corona_ts,order=c(2,1,2))
corona_ts_arima_forecast<-forecast(corona_ts_arima)
corona_ts_arima_forecast

plot(corona_ts_arima_forecast,
     xlab='Date Counts from April,01,2020',
     ylab='Expected Daily New Cases',
     main='ARIMA Model for the prediction from 9 June to 16 June')

# Monthly Forecast
# Total Cases...

auto.arima(total_cases_ts)
corona_ts_arima_total<-arima(total_cases_ts,order=c(3,2,1))
corona_ts_arima_total


corona_ts_arima_forecast_total<-forecast(corona_ts_arima_total,h=30)
corona_ts_arima_forecast_total

plot(corona_ts_arima_forecast_total,
     xlab='Date Counts from April,01,2020',
     ylab='Cumulative Cases',
     main='ARIMA Model for the prediction from 6 June to 6 July')
# Three Month Forcast
corona_ts_arima_forecast_total<-forecast(corona_ts_arima_total,h=90)
corona_ts_arima_forecast_total

plot(corona_ts_arima_forecast_total,
     xlab='Date Counts from April,01,2020',
     ylab='Cumulative Cases',
     main='ARIMA Model for the prediction from 6 June to 6 August')

```


### CONCLUSION FROM EXPONENTIAL MODEL

The prediction most likely to follow in between Lo95% to Hi95% Output
Date count starts from 2 June.....
```{r}
## One Week Daily New Cases
corona_ts_forecasts_exp2
plot(corona_ts_forecasts_exp2,main='Exponential Model Forecast for one Week(6June to 13 June)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Daily New Cases of Corona Infection')
## One Week Cumulative Cases
corona_ts_forecasts_total_cases_exp2
plot(corona_ts_forecasts_total_cases_exp2,main='Exponential Model Forecast for one Week(9June to 16 June)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')

#Total Cases for the month
corona_ts_forecasts_total_cases_exp2_month
plot(corona_ts_forecasts_total_cases_exp2_month,main='Exponential Model Forecast for one Week(6June to 6 August)',
     xlab = 'Date Count starts from April1,2020',
     ylab ='Cumulative Cases of Corona Infection')



```
